---
title: "Trends.Earth Usage Report"
author: "Automated run on Trends.Earth API server"
date: "`r format(Sys.time(),'%A, %B %d, %Y %H:%M:%S %Z')`"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(RPostgres)
library(ggplot2)
library(scales)
library(dplyr)
library(stringr)
library(lubridate)
library(foreach)
library(jsonlite)
library(geojsonsf)
library(forcats)
library(ggrepel)
library(RColorBrewer)
library(rnaturalearth)
library(rnaturalearthdata)
library(cowplot)
library(lwgeom)
library(logger)
library(ggnewscale)

world <- ne_countries(scale = "medium", returnclass = "sf")
crs_wintri <- "+proj=wintri +datum=WGS84 +no_defs +over"
world_wintri <- st_transform_proj(world, crs = crs_wintri)
grat_wintri <- 
  st_graticule(lat = c(-89.9, seq(-60, 60, 30), 89.9),
               lon = c(-179.9, seq(-135, 135, 45), 179.9)) %>%
  st_transform_proj(crs = crs_wintri)

period_start <- today() - years(1)
period_end <- now()

#readRenviron("rstudio.env")

con <- dbConnect(
    RPostgres::Postgres(),
    user=Sys.getenv("POSTGRES_USER"),
    password=Sys.getenv("POSTGRES_PASSWORD"),
    host=Sys.getenv('POSTGRES_HOST'),
    port=Sys.getenv("POSTGRES_PORT"),
    dbname=Sys.getenv("POSTGRES_DB_NAME")
)

script <- tbl(con, "script") %>% collect()
user <- tbl(con, "user") %>%
    collect() %>%
    mutate(email=abbreviate(email, 20))
jobs <- tbl(con, "execution") %>%
    select(script_id, user_id, status, start_date) %>%
    filter(start_date >= period_start,
           start_date <= period_end) %>%
    collect() %>%
    left_join(select(script, id, name, slug), by=c('script_id' = 'id')) %>%
    left_join(select(user, id, email), by=c('user_id' = 'id')) %>%
    filter(!str_detect(email, 'trends.earth')) %>%
    select(-email, -name) %>%
    mutate(task=str_remove(slug, '[0-9-]*$'),
           version=gsub('-', '.', str_extract(slug, '[0-9]+[0-9-]*$')),
           task=factor(task))

```

# Overall statistics

There have been `r jobs %>% filter(start_date >= (today() - years(1))) %>% summarize(n=n())` jobs run in the last year, `r jobs %>% filter(start_date >= (today() - months(1))) %>% summarize(n=n())` in the past month, `r jobs %>% filter(start_date >= (today() - days(7))) %>% summarize(n=n())` in the past week, and  `r jobs %>% filter(start_date >= (today() - days(1))) %>% summarize(n=n())` in the past day.

There have been `r user %>% filter(created_at >= (today() - years(1))) %>% summarize(n=n())` users added in the last year, `r user %>% filter(created_at >= (today() - months(1))) %>% summarize(n=n())` in the past month, `r user %>% filter(created_at >= (today() - days(7))) %>% summarize(n=n())` in the past week, and  `r user %>% filter(created_at >= (today() - days(1))) %>% summarize(n=n())` in the past day.

# Detailed statistics

## Statistics over the past 72 hours

### Hourly job submissions

```{r fig.width=12}
jobs %>%
    filter(start_date >= now(tzone='UTC') - hours(24*3)) %>%
    mutate(time=floor_date(start_date, unit='hours')) -> jobs_past_72_hours

# Make a dataframew with median jobs per hour so that the plot can also show the
# generally expected number of tasks per day/hour
data.frame(time=seq(min(jobs_past_72_hours$time),
                    max(jobs_past_72_hours$time), by='hour')) %>%
    mutate(wday=wday(time), hour=hour(time)) %>%
    left_join(
        group_by(jobs) %>%
            mutate(time=floor_date(start_date, unit='hours')) %>%
            group_by(time) %>%
            summarise(n=n()) %>%
            group_by(wday=wday(time), hour=hour(time)) %>%
            summarise(median_n=median(n))
    ) -> median_n

jobs_past_72_hours %>%
    group_by(time, status) %>%
    summarise(n=n()) %>%
    mutate(wday=wday(time), hour=hour(time)) %>%
    ggplot() +
    theme_minimal(base_size = 16) +
    geom_bar(aes(time, n, fill=status), stat='identity', alpha=.4) +
    guides(fill=guide_legend("Job status")) +
    new_scale_fill() +
    geom_bar(data=median_n, aes(time, median_n, fill='Median\nnumber of jobs\nover past year'), stat='identity', alpha=.1) +
    scale_fill_manual(name=element_blank(), values='black') +
    scale_x_datetime(date_labels="%H %Z\n%a\n%b %d", date_breaks="6 hours") +
    xlab('Time (UTC)') +
    ylab('Number of jobs submitted')
```

### Job status by task type
```{r fig.width=12}

# Top task types
jobs_past_72_hours %>%
    ggplot() +
    theme_minimal(base_size = 16) +
    geom_bar(aes(forcats::fct_infreq(task), fill=status)) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
    xlab('Task') +
    ylab('Number of jobs submitted')
```

### Ten top users and their tasks
```{r fig.width=12}
# Top 10 users and their top job slugs / statuses
jobs_past_72_hours %>%
    group_by(user_id) %>%
    summarise(n=n()) %>%
    slice_max(order_by=n, n=10) %>%
    left_join(jobs_past_72_hours) %>%
    left_join(user, by=c("user_id"="id")) %>%
    ggplot() +
    theme_minimal(base_size = 16) +
    geom_bar(aes(forcats::fct_infreq(email), fill=task)) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
    xlab('User') +
    ylab('Number of jobs submitted')
```

### Ten top users and their task statuses
```{r fig.width=12}
jobs_past_72_hours %>%
    group_by(user_id) %>%
    summarise(n=n()) %>%
    slice_max(order_by=n, n=10) %>%
    left_join(jobs_past_72_hours) %>%
    left_join(user, by=c("user_id"="id")) %>%
    ggplot() +
    theme_minimal(base_size = 16) +
    geom_bar(aes(forcats::fct_infreq(email), fill=status)) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
    xlab('User') +
    ylab('Number of jobs submitted')
```

### Code versions being used
```{r fig.width=12}
# Code versions being used
jobs_past_72_hours %>%
    ggplot() +
    theme_minimal(base_size = 16) +
    geom_bar(aes(forcats::fct_infreq(version), fill=task)) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
    xlab('Version') +
    ylab('Number of jobs submitted')

```

### Task status by code version
```{r fig.width=12}
# Code versions being used
jobs_past_72_hours %>%
    ggplot() +
    theme_minimal(base_size = 16) +
    geom_bar(aes(forcats::fct_infreq(version), fill=status)) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
    xlab('Version') +
    ylab('Number of jobs submitted')

```

## Statistics over the past month

```{r jobs_past_month_types, fig.width=12}
jobs %>%
    group_by(task, year=year(start_date), month=month(start_date), day=day(start_date)) %>%
    mutate(date=ymd(paste(year, month, day, sep='-'))) %>%
    filter(date >= today() - months(1)) -> jobs_last_month

# Top task types
jobs_last_month %>%
    ggplot() +
    theme_minimal(base_size = 16) +
    geom_bar(aes(forcats::fct_infreq(task), fill=status)) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
    xlab('Task') +
    ylab('Number of jobs submitted')

# Code versions being used
jobs_last_month %>%
    ggplot() +
    theme_minimal(base_size = 16) +
    geom_bar(aes(forcats::fct_infreq(version), fill=task)) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
    xlab('Version') +
    ylab('Number of jobs submitted')


# Top 10 users and their top job slugs / statuses
jobs_last_month %>%
    group_by(user_id) %>%
    summarise(n=n()) %>%
    slice_max(order_by=n, n=10) %>%
    left_join(jobs_last_month) %>%
    left_join(user, by=c("user_id"="id")) %>%
    ggplot() +
    theme_minimal(base_size = 16) +
    geom_bar(aes(forcats::fct_infreq(email), fill=task)) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
    xlab('User') +
    ylab('Number of jobs submitted')

jobs_last_month %>%
    group_by(user_id) %>%
    summarise(n=n()) %>%
    slice_max(order_by=n, n=10) %>%
    left_join(jobs_last_month) %>%
    left_join(user, by=c("user_id"="id")) %>%
    ggplot() +
    theme_minimal(base_size = 16) +
    geom_bar(aes(forcats::fct_infreq(email), fill=status)) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
    xlab('User') +
    ylab('Number of jobs submitted')

```

```{r jobs_past_month_number, fig.width=12}
jobs %>%
    group_by(task, year=year(start_date), month=month(start_date), day=day(start_date)) %>%
    summarise(n=n()) %>%
    mutate(date=ymd(paste(year, month, day, sep='-'))) %>%
    filter(date >= today() - months(1)) -> daily_tasks

daily_tasks %>%
    ggplot() +
    theme_minimal(base_size = 16) +
    geom_line(aes(date, n, colour=task, linetype=task)) +
    geom_point(aes(date, n, colour=task, shape=task)) +
    scale_colour_manual('Task', values = rep(c('#000000', '#3B69B0', '#D97D33', '#419753'), 3)) +
    scale_linetype_manual('Task', values = rep(1:4, 3)) +
    scale_shape_manual('Task', values = rep(1:4, each=3)) +
    xlab('Date') +
    ylab('Number of jobs submitted per day')

jobs %>%
    filter(start_date >= today() - months(1)) %>%
    ggplot() +
    theme_minimal(base_size = 16) +
    geom_bar(aes(start_date, fill=status)) +
    xlab('Date') +
    ylab('Number of jobs submitted per day')

user %>%
    filter(created_at >= period_start,
           created_at <= period_end) %>%
    group_by(year=year(created_at), month=month(created_at), day=day(created_at)) %>%
    summarise(n=n()) %>%
    mutate(date=ymd(paste(year, month, day, sep='-'))) %>%
    filter(date >= today() - months(1)) -> daily_users

daily_users %>% group_by(date) %>%
    summarise(n=sum(n)) %>%
    ggplot() +
    theme_minimal(base_size = 16) +
    geom_bar(aes(date, n), stat='identity') +
    xlab('Date') +
    ylab('Number of new users per day')
```

## Over the past year

```{r jobs_past_year, fig.width=12}
jobs %>% group_by(task, year=year(start_date), month=month(start_date)) %>%
    filter(status == 'FINISHED') %>%
    summarise(n=n()) %>%
    mutate(date=ymd(paste(year, month, '01', sep='-'))) -> monthly_tasks

monthly_tasks %>%
    ggplot() +
    theme_minimal(base_size = 16) +
    geom_line(aes(date, n, colour=task, linetype=task)) +
    geom_point(aes(date, n, colour=task, shape=task)) +
    scale_colour_manual('Task', values = rep(c('#000000', '#3B69B0', '#D97D33', '#419753'), 3)) +
    scale_linetype_manual('Task', values = rep(1:4, 3)) +
    scale_shape_manual('Task', values = rep(1:4, each=3)) +
    xlab('Date') +
    ylab('Number of jobs submitted per month')

monthly_tasks %>%
  group_by(date) %>%
  summarise(n=sum(n)) %>%
    ggplot() +
    theme_minimal(base_size = 16) +
    geom_bar(aes(date, n), stat='identity') +
    xlab('Date') +
    ylab('Number of jobs submitted per month')

user %>%
    filter(created_at >= (today() - months(12))) %>%
    group_by(year=year(created_at), month=month(created_at)) %>%
    summarise(n=n()) %>%
    mutate(date=ymd(paste(year, month, '01', sep='-'))) -> monthly_users

monthly_users %>%
    ggplot() +
    theme_minimal(base_size = 16) +
    geom_bar(aes(date, n), stat='identity') +
    xlab('Date') +
    ylab('Number of new users per month')

```


## User statistics

In the last year, users have joined from `r length(unique(user$country))` different countries.

```{r users_location_past_year, fig.width=12}
user_count <- nrow(user)
country_count <- length(unique(user$country))

user %>%
    group_by(country) %>%
    summarise(n=n()) %>%
    mutate(frac=n/sum(n)) %>%
    arrange(desc(n)) %>%
    top_n(20) %>%
    mutate(country=fct_reorder(country, n, .desc=TRUE)) %>%
    ggplot() +
    theme_minimal(base_size = 16) +
    geom_bar(aes(country, frac), stat='identity') +
    ylab('Fraction of all users') +
    ggtitle(paste0('Top 20 countries by number of users who joined between ', period_start,' to ', period_end, '\n(total users = ', user_count, ', total countries = ', country_count, ')')) +
    scale_y_continuous(labels=percent) +
    theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust=.5),
          axis.title.x=element_blank(),
          axis.ticks.x=element_blank(),
          panel.grid.minor.x = element_blank(),
          panel.grid.major.x = element_blank())

n_top <- 15
mycolors <- rev(colorRampPalette(brewer.pal(8, "YlGn")[1:6])(n_top+1))

user %>%
    group_by(country) %>%
    summarise(n=n()) %>%
    mutate(frac=n/sum(n)) %>%
    arrange(desc(n)) %>%
    slice_head(n=n_top) -> top_x_country_users

top_x_country_users %>%
    bind_rows(data.frame(country=paste0('Other\n(countries with < ', round(top_x_country_users$frac[n_top], 2)*100, '%)'),
                         n=user_count-sum(top_x_country_users$n),
                         frac=1-sum(top_x_country_users$frac))) %>%
    mutate(country=fct_reorder(country, n, .desc=TRUE)) %>%
    ggplot(aes(x="", y=frac, fill=country)) +
    theme_minimal(base_size = 16) +
    geom_bar(width=1, stat='identity', colour='grey') +
    coord_polar("y", start=3) +
    ylab('Fraction of all users') +
    labs(caption=paste0('Total users = ', user_count, ', Total number of countries = ', country_count)) +
    theme(axis.text = element_blank(),
          axis.title = element_blank(),
          panel.grid.minor = element_blank(),
          panel.grid.major = element_blank(),
          plot.caption = element_text(hjust=0.5, size=rel(1)),
          legend.position='none') +
    geom_text_repel(aes(label = paste0(round(frac*100), '%', ' - ', country)),
              position = position_stack(vjust = 0.5), size=5)


group_by(user, country) %>% summarise(n=n()) -> user_countries

# Fix some names so our users data matches with the R boundary data
user_countries$country[user_countries$country == 'United States'] <- 'United States of America'
user_countries$country[user_countries$country == 'Swaziland'] <- 'eSwatini'
user_countries$country[user_countries$country == 'Macedonia'] <- 'North Macedonia'

left_join(world_wintri, user_countries, by=c('name'='country')) %>%
    ggplot() +
    geom_sf(aes(fill=n)) +
    geom_sf(data=grat_wintri, color = "gray90", size = 0.05/.pt) + 
    coord_sf(datum=st_crs("+proj=wintri +datum=WGS84 +no_defs +over")) +
    labs(fill='Number of\nnew users') +
    theme_map()

```
